---
title: "HW R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Calculator

```{r}
5 + 3
5 * 3
5 / 3
5 ^ 3
5 %% 3
5 %/% 3
```

## Sequenece

```{r}
1:10
seq(1, 10, 2)
seq(1, 10, length.out = 6)
```

## Basic features in R
We tell R to make the assignment using an arrow “<-”. R also supports
using “=” in most circumstances. However, “<-” is recommended as it
makes clear that we are requesting an assignment, rather than stating a
relation.

```{r}
a <- 3 ^ 5
b = 3 ^ 5
a
```

## Calculation by the set values

```{r}
a <- 1000
b <- 4
a * b - a
```
If you don’t assign a value to a name of variable, then you will get the error
message

## More on calculation
#### Absolute value: |x|

```{r}
abs(-2)
```
#### Exponential & logarithm: exp(x), log(x)
```{r}
exp(10); log(2); log2(2); log(10); log10(10);
```

## Other functions
“ˆ” or "**" means the degree of the power.

```{r}
2 ** 3; 2^3;
2^0.5; sqrt(2);

```

## Approximated values 

```{r}
round(125.78253)
round(125.78253, digits = 3)
round(125.78253, digits = -2)
round(83524587, digits = -3); round(83524587, digits = -5)
signif(125.78253, digits = 2)
floor(125.78253)
ceiling(125.78253)
trunc(125.78253)
```

## Descriptive statistics

```{r}
x <- 1:25
mean(x)
var(x)
median(x)
quantile(x, c(0.1, 0.5, 0.7))
```

## Creat a vector
We use c() to be a collection of needed values.

```{r}
x <- c(1, 3, 5, 7, 9, 11)
y <- c(2, 4, 6, 8, 10, 12)
x
y
```

## Easy way to get the summary of a set

```{r}
x <- c(1:25, NA)
summary(x)
```

## Figures of a set

```{r}
x <- c(1:25, NA)
par(mfrow = c(1, 2))
boxplot(x)
hist(x)

```

## Some common functions to vectors
Creat two vectors and add one vector to another.
```{r}
vt1 <- c(2:4)
vt2 <- c(3:6)
vt2 <- c(vt2, vt1)
vt2
length(vt1)
length(vt2)
vt1[3]
vt2[c(3,5,7)]
vt1[-2]
vt2[-c(3:6)]
```

## Vector arithmetic

```{r}
x <- c(3, 6, 9)
x * 3
x - 3
x ^ 3
y1 <- c(3, 2, 4);
x^y1
```

## Simple patterned vectors
Use “seq()” and “rep()” to construct a vector with similar patterns.
```{r}
rep(3, 12)
rep(seq(2, 10, 2), 2)
rep(c(1, 4), c(3, 2))
rep(seq(2, 10, 2), rep(3, 5))
```

##  Construct a matrix by row
A matrix is a generalization containing multiple vectors.
```{r}
a <- 1:5; b <- 6:10;
c <- 11:15; d <- 16:20
M1 <- rbind(a, b, c, d)
M1

```

## Construct a matrix by column
```{r}
M2 <- cbind(a, b, c, d)
M2
dim(M1)
dim(M2)

```

## Constrict a matrix by a long vector

```{r}
X <- 1:20
matrix(X, nrow = 4, ncol = 5)
```

## Features for a matrix

```{r}
a <- 1:5; b <- 6:10;
c <- 11:15; d <- 16:20
M1 <- rbind(a, b, c, d)
colnames(M1); rownames(M1)
colnames(M1) <- c("V1", "V2", "V3", "V4", "V5")
M1

```

## Extract elements from a matrix

```{r}
a <- 1:5; b <- 6:10;
c <- 11:15; d <- 16:20
M1 <- rbind(a, b, c, d)
M1[2, 4]
M1[c(2, 4), 3]; M1[c(2, 4), c(3, 5)];
M1[5]
```

## Elements in the vector

```{r}
b <- c(3, 3, 3, 3, 5, 5, 1, 2, 10, 10, 8)
unique(b)
table(b)

```

## Set Operations

```{r}
B <- c(3, 1, 5, 4, 7); C <- c(23, 1, 44, 5, 3)
union(B, C)
intersect(B, C)
setdiff(B, C)
setequal(B, C)
```

## Sorting and ordering of a vector

```{r}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
sort(a)
sort(a, decreasing = TRUE)
inc.order <- order(a)
dec.order <- order(a, decreasing = TRUE)
inc.order
dec.order
a[inc.order]
a[dec.order]
```

## Change values in a vector

```{r}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
a[5] <- -3
a[8] <- -1
a
```

## Character vectors
* Character vector: the vector is made up of strings of characters.
* All elements of a vector must be of the same type.
```{r}
colors <- c("red", "yellow", "BLUE")
more.colors <- c(colors, "green", "cyan")
more.colors
```

## Basic operations for character vectors
-To take substrings, use “substr( )”.
```{r}
colors <- c("red", "yellow", "blue")
substr(colors, 1, 3)

```
-To search a string from a vector, use “grep()”, which is “Global Regular
Expression Print”. It returns the locations of the desired patterns.
```{r}
colors <- c("red", "yellow", "blue")
grep("y", more.colors)
grep("e", more.colors)
```

## Basic operations for character vectors
-To building up strings, use “paste()”.
```{r}
colors <- c("red", "yellow", "blue")
paste(colors, "line") 
paste(colors, "line", sep = "" ) 
paste("I like", colors, collapse = "; ")
```

## Factors
factor(): factor(x = character())
```{r}
colors <- c("red", "yellow", "blue")
f.c <- factor(colors)
f.c
levels(f.c)
as.integer(f.c)
```

## Logical number
* Number: Compare 3 and 5
* “==”: equal
* “!=”: not equal
```{r}
c(3 < 5, 3 > 5)
c(3 <= 5, 3 >= 5)
3 == 5
3 != 5
```

## Boolean operations

```{r}
A <- c(T, T, F, T, F); B <- c(3, 1, 5, 4, 7)
B[A]
B[!A]
```

#### How many TRUE (or FALSE) in the vector?
```{r}
A <- c(T, T, F, T, F)
sum(A)
sum(!A)
length(A == TRUE)
```

## Logical vector
#### Compare the same location of elements individually.
```{r}
B <- c(3, 1, 5, 4, 7); C <- c(23, 1, 44, 5, 3)
B == C
D <- c(2, 5 ,6, 10)
B == D
```

## “is.” series
#### “is.XYZ()” is any comparison to the type “XYZ” of an object.
* is.null()
* is.na()
* is.nan()
* is.infinite()
* is.character()
* is.integer()
* is.numeric()
```{r}
is.null(c())
c(is.na(NA), is.na(2))
x1 <- 1:5
x2 <- c(1:3, 2.2)
x3 <- colors
c(is.integer(x1), is.numeric(x1), is.double(x1))
c(is.integer(x2), is.numeric(x2), is.double(x2))
c(is.character(x3), is.numeric(x3))
```

#### Examine the type of a vector
Use “class( )” to identify the type of a vector.
```{r}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
b <- c("red", "yellow", "blue")
c <- c(TRUE, FALSE)
c(class(a), class(b), class(c))
c(class(2), class(2.0), class(as.integer(2)))
```

## For() loop
### When to use?
* See patterns of behaviour, not just a single instance.
* The for() statement allows one to specify that a certain operation
should be repeated a fixed number of times.
```{r}
vec.X.v1 <- c()
Xn <- 3
for(n in 1:7){
Xn <- 2*Xn + 3
vec.X.v1 <- c(vec.X.v1, Xn)
}
vec.X.v1
```

### Evaluate n!
Try n = 10
```{r}
n <- 10
ans <- 1
for(j in n:1){
ans <- ans * j
}
ans
factorial(n)
```

### Random number generator
Random samples from [0, 1] by the linear congruntial generator.
xn = (axn−1 + c) mod m.
Let x0 = 3, a = 5, c = 7, and m = 501. Find xi
, where i = 1, . . . , n. What
are the values after all in Xn devided by m?
```{r}
x0 <- 3; a <- 5; c <- 7; m <- 501
X <- rep(NA, 10)
for(j in 1:10){
x0 <- (a*x0+c) %% m
X[j] <- x0
}
X
round(X/m, 2)
```

## For loop in matrix
Given a matrix
X =
$\left[
\begin{array}{cc}
 3 & 2 & 4 & 5 & 3\\
 2 & 6 & 8 & 10 & 4\\
 1 & 4 & 6 & 8 & 3\\
 2 & 1 & 2 & 6 & 4\\
\end{array}
\right]$ ,
evaluate the mean for each row.
```{r}
A <- matrix(c(3, 2, 1, 2, 2, 6, 4, 1,
4, 8, 6, 2, 5, 10, 8, 6,
3, 4, 3, 4), 4, 5)
nr <- dim(A)[1]
m <- c()
for(j in 1:nr){
m <- c(m, mean(A[j, ]))
}
m
```

## if()
* This statement causes a set of commands to be invoked if condition
evaluates to TRUE.
* The else part is optional, and provides an alternative set of commands
which are to be invoked in case the logical variable is FALSE.
Given C = {3, 5, NA, 12, 0.2, NA, −2}, find the locations of NA.
```{r}
C <- c(3, 4, NA, 12, 0.2, NA, -2)
na.loc <- c() 
for(i in 1:7){
if(is.na(C[i])) na.loc <- c(na.loc, i)
}
na.loc
```

## while()
* while(): Need the initial, condition of stopping, and iteration or replacement
```{r}
j <- 1
while(j <= 8){
print(j)
j <- j + 1
}
```

### Fibonacci numbers (1)

List all Fibonacci numbers less than 100.
Can you know the location that a Fibonacci number is small than 100 in advance?
Use for() or while()?
```{r}
Fib1 <- 1
Fib2 <- 1
Fibonacci <- c(Fib1)
while (Fib2 < 100) {
Fibonacci <- c(Fibonacci, Fib2)
Fib1 <- Fibonacci[length(Fibonacci) - 1]
Fib2 <- Fib1 + Fibonacci[length(Fibonacci)]
}
Fibonacci
```

### Flip a coin
Let 0 be a tail and 1 be a head. Count the total number of flipping until each success occurs.
```{r}
coin <- c(0, 0, 0, 0, 0, 0, 1, 0, 1, 0, 1, 0,
1, 1, 0, 1, 1, 0, 0, 0, 1, 0, 1, 0,
0, 0, 0, 0, 0, 0, 1)
N <- c(); j <- 1
while(j <= length(coin)){
n.fail <- 0
while(coin[j] != 1){
n.fail <- n.fail + 1
j <- j + 1
}
N <- c(N, n.fail + 1)
j <- j + 1
}
N
```

## break & repeat
* break
if (condition) break
The break statement causes the loop to terminate immediately. break statements can also be used in for() and while() loops.
* repeat
Use “repeat” to repeat until we execute a break statement.
repeat { statements }
This causes the statements to be repeated endlessly.

```{r}
Fib1 <- 1
Fib2 <- 1
Fibonacci <- c(Fib1)
repeat
{
if(Fib2 < 100){
Fibonacci <- c(Fibonacci, Fib2)
Fib1 <- Fibonacci[length(Fibonacci) - 1]
Fib2 <- Fib1 + Fibonacci[length(Fibonacci)]
}else{
break
}
}
Fibonacci
```

## source( )
1. Creat a new script for your function or codes.
2. Save the script named as “XXX.R” and use **source( )** to run the code.
3. Remember to use **print( )** or **cat( )** so that your final answer will be shown.
#### Example:
```{r}
x <- c(0.8933, 0.2344, 0.0764, 0.7736)
x.percent <- round(100*x, digits = 1)
x.show <- paste(x.percent, "%", sep = "")
print(x.show)
cat(x.show)
```

## function( )
Use **function( )** when you only have one main code for evaluation and
multiple possible input, like
$$f (x) = 3x + 2, x = 2, 3, 4$$
f means a name of a function of input x when x = 2, 3, 4.
```{r}
f <- function(x){
3*x + 2
}
f(2); f(3); f(4)
```

### return( ) in the function( )
You must tell what will be return when you use the function by **return( )** or **invisible( )**. If you don’t put **return( )** at the end of the function, it will treat the last line as the result automatically.
```{r}
f <- function(x){
return(3*x + 2)
}
f(2)
```

### Remark for the functions
* The name of a function is also a type of an object.
* If there are a vector and a function with the same name, then the type of the object will be the last one.
Example
```{r}
objA <- function(x){
return(2*x)
}
objA(4)
```

##  list( )
**list( )** is a collection of multiple types of objects, like vectors, characters, numbers,lists,. . . .
#### Example:
```{r}
X <- list(a = 3, b = c(1,2,3,4,5), c = c("A", "B"))
X
```

## Use function( ) and list( )
Try to use list( ) to return more than one output. Example: perfect number
```{r}
perfect.number <- function(a, b){
k <- a
res <- c()
ans <- list()
while(k < b){
fac <- c()
for(j in 1:k){
if((k %% j) == 0) fac <- c(fac, j)
}
if(k == sum(fac[-length(fac)])) {
res <- c(res, k)
ans[[paste(k)]] <- fac[-length(fac)]
}
k <- k+1
}
ans$number <- res
return(ans)
}
Ans <- perfect.number(1, 100)
Ans
```

## data.frame( )
* Most data sets are stored in R as data frames.
* These are like matrices, but with the columns having their own names.
* Columns can be of different types from each other.
Use the **data.frame()** to construct data frames from vectors:
```{r}
colors <- c("red", "yellow", "blue")
numbers <- c(1, 2, 3)
colors.and.numbers <- data.frame(colors, numbers,
more.numbers = c(4, 5, 6))
colors.and.numbers
```

### Types in a data frame
Use **str()** to examine the types of columns in the data frame.
```{r}
str(colors.and.numbers)
```

## Statistical Figures
### Graphics for descriptive statistics
#### Given an arbitray vector x:
1. Histogram: hist(x)
2. Boxplot: boxplot(x)
3. Density plot: plot(density(x))
```{r}
x.norm <- rnorm(100, mean=0, sd=1)
par(mfrow = c(1, 3))
hist(x.norm)
boxplot(x.norm)
plot(density(x.norm))
```

#### For discrete variables x:
1. Barchart: barplot(x)
2. Piechart: pie(x)
```{r}
x <- sample(c(1:8), 100, replace=TRUE, prob=c(8:1))
x <- table(x)
names(x) <- c("A", "B", "C", "D", "E", "F", "G", "H")
par(mfrow = c(1, 2))
barplot(x)
pie(x)
```

#### Scatter plot is for the correlation between two variables.
Use **plot(x, y)**
```{r}
x <- rnorm(100)
y <- 3*x + rnorm(100, 0, 1)
plot(x, y)
cor(x, y)
```

### General plot function: plot(. . . )
plot(x, y,
type = "p", 
pch = , 
main = NULL, 
lty = , 
lwd = , 
cex = ,
bty="o", 
col = , 
xlim = NULL, ylim = NULL,
sub = NULL, 
xlab = NULL, ylab = NULL, 
cex.main = 
)
### Add points or lines to an existing plot
* **points(. . . )**: add points
* **lines(. . . )**: add lines
* **abline(. . . )**: add straight lines
$$abline(a = ..., b = ..., h = ..., v = ..., reg = ..., coef = ...)$$
$$y = a + bx$$
```{r}
t <- 1:20
y <- rnorm(20)
plot(t, y, type = "b")
abline(h = 0)
abline(v = 15, lty = 2, col = 8, lwd = 2)
points(t[15], y[15], pch = 19, col = 2)

```

### Add text to an existing plot
```{r}
x <- rnorm(50)
y <- 3*x + rnorm(50, 0, 1)
y1 <- 5*x + rnorm(50, 0, 2)
plot(x, y, type = "p", pch = 19)
abline(a = 0, b = 3)
points(x, y1, col = 4, pch = 20)
abline(a = 0, b = 5, col = 4)
text(-1.5, 2, "A", cex = 5)
text(1, -5, "B", cex = 5, col = 2)
legend("topleft", c("Y", "Y1"), col = c(1,4), lty = c(1, 1),
cex = 2)

```

